# Regression modeling and inference

## Notes

<embed src="lectures/notes-regression-modeling-and-inference.pdf" type="application/pdf" width="100%" height="1000px"></embed>

## Lab 1

```{r}
#| label: load-prereqs
#| echo: false
#| message: false

# knitr options
source("_common.R")

# needed if run stan so that it compiles without any errors
Sys.setenv(
  SDKROOT = "/Library/Developer/CommandLineTools/SDKs/MacOSX.sdk",
  CXXFLAGS = "-stdlib=libc++",
  CXX11FLAGS = "-stdlib=libc++",
  CXX14FLAGS = "-stdlib=libc++",
  CXX17FLAGS = "-stdlib=libc++"
)

```

### Normal models with unknown mean and variance

- Data:

$$
Y_1, \ldots, Y_n \overset{iid}\sim N(\mu,\sigma^2)
$$

$$
p(\mathbf{y} \mid \mu, \sigma^2) \propto (\sigma^2)^{-n/2}
\exp\left(-\frac{\sum_{i=1}^n (Y_i - \mu)^2}{2\sigma^2}\right)
$$

- Here are the BMI measurements of patients in one clinic:

```{r}

y <- c(18.3,19.4,19.7,23.4,30.9,21.7,21.7,21.4,26.3, 25.3)
y_bar <- mean(y)
n <- length(y)
s2 <- var(y)

```

#### What if we use noninformative priors?

- Prior:

  $$
  p(\mu,\sigma^2) \propto \frac{1}{\sigma^2}
  $$

- Posteriors:

  - Conditional posterior of $\mu$ given $\sigma^2$ and $\mathbf{y}$:
  
    $$
    \mu \mid \sigma^2, \mathbf{y} \sim \text{N}\left(\bar{y}, \frac{\sigma^2}{n}\right)
    $$
  
  - Marginal posterior of $\sigma^2$ given $\mathbf{y}$:
  
      $$
      \sigma^2 \mid \mathbf{y} \sim \text{Inv-Gamma}\left(\frac{n-1}{2}, \frac{(n-1)s^2}{2}\right)
      $$

- Simulation takes two steps:
  
  1. Simulation of $\sigma^2$ from the inverse gamma distribution (note that an inverse gamma random variable can be expressed as the reciprocal of a gamma random variable with the same shape and rate parameters. Alternatively, you can use `invgamma::rigamma()`).
  
  2. Simulation of $\mu$ from the normal distribution

```{r}

# initialize items
m <- 1000

# sample from marginal posterior for sigma^2 given y
post_sigma2 <- 1 / rgamma(n = m, shape = (n - 1) / 2, rate = (n - 1) * s2 / 2)

# sample from conditional posterior for mu given sigma^2 and y
post_mu <- rnorm(n = m, mean = y_bar, sd = sqrt(post_sigma2 / n))

# plot the distributions
hist(post_sigma2)
hist(post_mu)

# point estimates
c("med sigma2" = median(post_sigma2), "mean sigma2" = mean(post_sigma2), "med mu" = median(post_mu), "mean mu" = mean(post_mu))

# 95% equal tail interval estimates
quantile(post_sigma2, probs = c(0.025,0.975))
quantile(post_mu, probs = c(0.025,0.975))

# 95% HPD interval estimates
TeachingDemos::emp.hpd(post_sigma2)
TeachingDemos::emp.hpd(post_mu)

```

Now we can plot the joint posterior density for $\mu$ and $\sigma^2$.

```{r}

# initialize items
# -> values of parameters for which you want to draw the plot with
# -> results holder
x_mu <- seq(from = 20, to = 26, length = 600)
x_sigma2 <- seq(from = 3, to = 25, length = 600)
post_joint <- matrix(data = 0, nrow = 600, ncol = 600)

# calculate joint posterior density
for (i in 1:600) {
  
  # -> P(mu, sigma^2 | y) = P(mu | sigma^2, y) * P(sigma^2 | y)
  post_joint[,i] = invgamma::dinvgamma(x = x_sigma2[i], shape = (n - 1) / 2, rate = (n - 1) * s2 / 2) * dnorm(x = x_mu, mean = y_bar, sd = sqrt(x_sigma2[i] / n))

}

# draw plots
contour(x = x_mu, y = x_sigma2, z = post_joint)
image(x = x_mu, y = x_sigma2, z = post_joint, xlab = "mu", ylab = "sigma^2")

```

```{r}
#| eval: false

# NOT RUN
persp(x = x_mu, y = x_sigma2, z = post_joint, xlab = "mu", ylab = "sigma^2")

```

#### What if we use conjugate priors?

- Prior:

  - Marginal prior of $\sigma^2$:
  
    $$
    \sigma^2 \sim \text{Inv-Gamma}(\alpha,\beta)
    $$
    
  - Conditional prior of $\mu$ given $\sigma^2$:
  
    $$
    \mu \mid \sigma^2 \sim N\left(\mu_0, \frac{\sigma^2}{\kappa_0}\right),
    $$
    
  where $\kappa$ reflects our confidence in the prior.
  

- Posteriors:

  - Conditional posterior of $\mu$ given $\sigma^2$ and $\mathbf{y}$:
  
    $$
    \mu \mid \sigma^2,\mathbf{y} \sim \text{N}(\mu_n,\sigma_n^2),
    $$
    
  where $\mu_n = \frac{n\bar{y} + \kappa_0 \mu_0}{n + \kappa_0}, \quad \sigma_n^2 = \frac{\sigma^2}{n + \kappa_0}$
  
  - Marginal posterior of $\sigma^2$ given $\mathbf{y}$:
  
      $$
      \sigma^2 \mid \mathbf{y} \sim \text{Inv-Gamma}\left(
\frac{n}{2}+\alpha, \frac{(n-1)s^2}{2}+\frac{n\kappa_0}{2(n+\kappa_0)}(\mu_0-\bar{y})^2+\beta\right)
      $$
      
Use the same two-step process to simulate from the posterior distribution using $µ_0 = 20$, $\kappa_0 = 0.01$, $\alpha = 0.1$ and $\beta = 10$. Report the marginal posterior median, mean, credible intervals for $\mu$ and $\sigma^2$. Draw the posterior distribution of $(\mu,\sigma^2)$.

```{r}

# initialize items
m <- 1000
alpha_0 <- 0.1
beta <- 10
mu_0 <- 20
k_0 <- 0.01

# sample from marginal posterior for sigma^2 given y
alpha_post <- n / 2 + alpha_0
beta_post <- (n - 1) * s2 / 2 + n * k_0 / (2 * (k_0 + n)) * (mu_0 - y_bar)^2 + beta
post_sigma2_conj <- invgamma::rinvgamma(n = m, shape = alpha_post, rate = beta_post)

# calculate posterior distribution mean
mu_n <- (n * y_bar + k_0 * mu_0) / (n + k_0)

# sample from conditional posterior for mu given sigma^2 and y
post_mu_conj <- rnorm(n = m, mean = mu_n, sd = sqrt(post_sigma2_conj / (n + k_0)))

# point estimates
c("med sigma2" = median(post_sigma2_conj), "mean sigma2" = mean(post_sigma2_conj), "med mu" = median(post_mu_conj), "mean mu" = mean(post_mu_conj))

# 95% equal tail interval estimates
quantile(post_sigma2_conj, probs = c(0.025,0.975))
quantile(post_mu_conj, probs = c(0.025,0.975))

# 95% HPD interval estimates
TeachingDemos::emp.hpd(post_sigma2_conj, conf = 0.95)
TeachingDemos::emp.hpd(post_mu_conj, conf = 0.95)

```

```{r}

# initialize items
# -> values of parameters for which you want to draw the plot with
# -> results holder
x_mu <- seq(from = 20, to = 26, length = 600)
x_sigma2 <- seq(from = 3, to = 25, length = 600)
post_joint_conj <- matrix(data = 0, nrow = 600, ncol = 600)

# calculate joint posterior density
for (i in 1:600) {
  
  # -> P(mu, sigma^2 | y) = P(mu | sigma^2, y) * P(sigma^2 | y)
  post_joint_conj[,i] = invgamma::dinvgamma(x = x_sigma2[i], shape = alpha_post, rate = beta_post) * dnorm(x = x_mu, mean = y_bar, sd = sqrt(x_sigma2[i] / n))

}

# draw plot
contour(x = x_mu, y = x_sigma2, z = post_joint_conj)

```

### Multinomial models with conjugate priors

#### Setup

Experiment:

- Fixed number of $n$ trials.
- Each trial is an independent event.
- The same $J$ outcomes are possible in each trial.
- Probability of each outcome to occur is constant for each trial.
- The multivariate response $Y$ equals the number of diﬀerent outcomes among these trials.

Example:

30 aircrafts that have been in service for 10 years were examined in a safety test. Among them,

- 19 have no wing cracks (Status 1)
- 9 have detectable wing cracks (Status 2)
- 2 have critical wing cracks (Status 3)

```{r}

data_aircraft <- c(1,3,1,1,1,1,1,1,1,1,2,1,1,1,2,2,2,2,2,1,2,3,2,1,1,1,1,2,1,1)
table(data_aircraft)

# save summary data
y <- as.numeric(table(data_aircraft))

```

Could you estimate the probability of aircrafts having status 1, 2 and 3 after serving for 10 years?

- Data: $Y_1, \ldots, Y_J \sim \text{Multinomial}(n,\theta_1,\ldots,\theta_J) \quad \text{where } n = Y_1 + \ldots + Y_n,n=30,J=3$
- Conjugate prior: $(\theta_1,\ldots,\theta_J) \sim \text{Dirichlet}(\alpha_1,\ldots,\alpha_J)$

How can we implement the Bayesian model to obtain the posterior samples of $\theta_1,\ldots,\theta_J$?

#### One prior

Just using some selected values for $\mathbf{\alpha} = (3,2,1)$.

IMPORTANT NOTE: Because of the conjugate prior, we know that the posteriors for $\mathbf{\theta}$ will also follow Dirichlet distribution, just with updated parameters.

```{r}

# set prior parameters
alphas <- c(3,2,1)

# calculate posterior parameters
post_params <- y + alphas

# calculate exact posterior mean for each of theta1, theta2, theta3
(post_theta_means <- post_params / sum(post_params))

# Monte Carlo simulation
post_thetas <- gtools::rdirichlet(n = 10000, alpha = post_params)

# approximate posterior means for each of theta1, theta2, theta3
post_thetas %>% 
  data.frame %>% 
  map_dbl(\(col) mean(col))

# approximate posterior medians and 95% credible intervals for each of theta1, theta2, theta3
post_thetas %>% 
  data.frame %>% 
  map(\(col) quantile(col, probs = c(0.025,0.50,0.975)))

```

#### Uniform prior

```{r}

# set prior parameters
alphas_unif <- c(1,1,1)

# calculate posterior parameters
post_params_unif <- y + alphas

# calculate exact posterior mean for each of theta1, theta2, theta3
(post_theta_means_unif <- post_params_unif / sum(post_params_unif))

# Monte Carlo simulation
post_thetas_unif <- gtools::rdirichlet(n = 10000, alpha = post_params_unif)

# approximate posterior means for each of theta1, theta2, theta3
post_thetas_unif %>% 
  data.frame %>% 
  map_dbl(\(col) mean(col))

# approximate posterior medians and 95% credible intervals for each of theta1, theta2, theta3
post_thetas_unif %>% 
  data.frame %>% 
  map(\(col) quantile(col, probs = c(0.025,0.50,0.975)))

```

#### Strong prior

```{r}

# set prior parameters
# NOTE: increasing alpha's in prior distribution gives stronger prior input
alphas_strong <- c(1,1,1)

# calculate posterior parameters
post_params_strong <- y + alphas

# calculate exact posterior mean for each of theta1, theta2, theta3
(post_theta_means_strong <- post_params_strong / sum(post_params_strong))

# Monte Carlo simulation
post_thetas_strong <- gtools::rdirichlet(n = 10000, alpha = post_params_strong)

# approximate posterior means for each of theta1, theta2, theta3
post_thetas_strong %>% 
  data.frame %>% 
  map_dbl(\(col) mean(col))

# approximate posterior medians and 95% credible intervals for each of theta1, theta2, theta3
post_thetas_strong %>% 
  data.frame %>% 
  map(\(col) quantile(col, probs = c(0.025,0.50,0.975)))

```

## Lab 2

### The coin example

- Experiment; Suppose that a coin is tossed 20 times, but Chloe only recorded the first 19 outcomes:
$\{1,1,1,1,1,0,1,1,1,1,0,1,1,0,1,1,0,0,1\}$
- Data: 

  - $X_1, X_2, \ldots, X_{19}$ taking on values of 1 (head) or 0 (tail), $X_{20}$ is missing.
  - $X_i \sim \text{Bernoulli}(\theta)$

- Parameter of interest: $\theta$.
- We can use Gibbs sampler to find posterior distribution of both $\theta$ and $X_{20}$!

```{r}

# obsersved data
# Observed data
x <- c(1,1,1,1,1,0,1,1,1,1,0,1,1,0,1,1,0,0,1)
y <- sum(x)

# initialize items
m <- 10000
m_burn_in <- 5000
post_theta <- c()
post_x20 <- c()

# Gibbs sampler!
# -> set initial values of p and X20
post_theta[1] <- 0.5
post_x20[1] <- 1

# add values to the chain
for (i in 2:m){
  
  # derived in previous notes
  post_theta[i] <- rbeta(n = 1, shape1 = y + post_x20[i-1], shape2 = 20 - y - post_x20[i-1] + 1)
  post_x20[i] <- rbinom(n = 1, size = 1, prob = post_theta[i])

}

# time series plot of the Markov Chain path for "p"
plot(post_theta, type = 'l')

# display and summarize kept values
hist(post_theta[(m_burn_in + 1):m])
mean(post_theta[(m_burn_in + 1):m])
sd(post_theta[(m_burn_in + 1):m])
table(post_x20[(m_burn_in + 1):m])

```

### Normal models with noninformative priors

Consider a normal models with noninformative priors. $Y_1, \ldots, Y_n \sim \text{N}(\mu, \sigma^2)$. The data set is simulated from the below code.

```{r}

y <- rnorm(100,20, 5)

```

Use Rstan to build the model and obtain the posterior distribution of $\mu$ and $\sigma^2$. Draw histograms for each parameter. Obtain the posterior predictive distribution for a future observation. Draw a histogram.

```stan
data {
  int<lower=0> n;
  real y[n];
}
parameters {
  real mu;
  real<lower=0> sigma2;
}
model {
  y ~ normal(mu, sqrt(sigma2));
  target += -log(sigma2);
}
generated quantities {
  real y_pred;
  y_pred = normal_rng(mu, sqrt(sigma2));
}
```

```{r}
#| cache: true

# specify stand code
stan_code <- "
data {
  int<lower=0> n;
  real y[n];
}
parameters {
  real mu;
  real<lower=0> sigma2;
}
model {
  y ~ normal(mu, sqrt(sigma2));
  target += -log(sigma2);
}
generated quantities {
  real y_pred;
  y_pred = normal_rng(mu, sqrt(sigma2));
}
"

# save model, provide data and run
stan_mod <- rstan::stan_model(model_code = stan_code)
data_stan <- list(n = length(y),
                  y = y)
stan_fit <- rstan::sampling(object = stan_mod, data = data_stan,
                            seed = 1, chains = 2, warmup = 5000, iter = 10000)

# view results
rstan::summary(stan_fit)
print(stan_fit, digits = 2)

# save and visualize results
list_draws <- rstan::extract(stan_fit)
hist(list_draws$mu)
hist(list_draws$sigma2)
hist(list_draws$y_pred)
mean(list_draws$mu)
mean(list_draws$sigma2)

```