# Regression modeling and inference

## Notes

<embed src="lectures/notes-regression-modeling-and-inference.pdf" type="application/pdf" width="100%" height="1000px"></embed>

## Lab

```{r}
#| label: load-prereqs
#| echo: false
#| message: false

# knitr options
source("_common.R")

# needed if run stan so that it compiles without any errors
Sys.setenv(
  SDKROOT = "/Library/Developer/CommandLineTools/SDKs/MacOSX.sdk",
  CXXFLAGS = "-stdlib=libc++",
  CXX11FLAGS = "-stdlib=libc++",
  CXX14FLAGS = "-stdlib=libc++",
  CXX17FLAGS = "-stdlib=libc++"
)

```

### Bayesian linear regression models: Analysis with noninformative priors

Normal model, noninformative prior.

Data:
- $y$: change in maximal oxygen uptake of kids (positive number = improvement)
- $X_1$: 0 = running group, 1 = aerobics group
- $X_2$: ages of subjects
- $X_3$: interaction between group and age

```{r}

y <- c(-0.87,-10.74,-3.27,-1.97,7.50,-7.25, 17.05,4.96,10.40,11.05,0.26,2.51)
x1 <- c(0,0,0,0,0,0,1,1,1,1,1,1)
x2 <- c(23,22,22,25,27,20,31,23,27,28,22,24)
x3 <- x1 * x2
X <- cbind(x0 = rep(1, length(y)), x1, x2, x3)

```

#### Using independent samples

Now we can simulate from the posterior.

```{r}

# solve for posterior parameters
b_hat <- solve(t(X) %*% X) %*% t(X) %*% y
post_alpha <- (nrow(X) - ncol(X)) / 2
post_gamma <- t(y - X %*% b_hat) %*% (y- X %*% b_hat) / 2

# initialize items
m <- 10000
post_beta <- matrix(data = 0, nrow = m , ncol = 4)
colnames(post_beta) <- paste0("b",0:3)

# generate samples
# -> initialize needed items
post_sigma2 <- 1 / rgamma(n = m, shape = post_alpha, rate = post_gamma)
X_t_Xinv <- solve(t(X) %*% X)
for(i in 1:m) {
  
  post_beta[i,] <- MASS::mvrnorm(1, b_hat, post_sigma2[i] * X_t_Xinv)

}

# summarize the posteriors
summary(post_beta)
summary(post_sigma2)

# find quantities of interest
mean(post_beta[,2] > 25)
quantile(post_beta[,2], probs = c(0.025,0.975))
quantile(post_beta[,3], probs = c(0.025,0.975))
quantile(post_beta[,4], probs = c(0.025,0.975))

# posterior prediction
# -> initialize items
x1_new <- c(0,1)
x2_new <- rep(mean(x2), 2)
x3_new <- x1_new * x2_new
X_new <- cbind(x0_new = 1, x1_new, x2_new, x3_new)
y_pred <- matrix(0, m, 2)

# generate predictions
for(i in 1:m) {

  mean_i <- X_new %*% post_beta[i,]
  y_pred[i,1] <- rnorm(n = 1, mean = mean_i[1], sd = sqrt(post_sigma2[i]))
  y_pred[i,2] <- rnorm(n = 1 ,mean = mean_i[2], sd = sqrt(post_sigma2[i]))

}

# find quantity of interest
quantile(y_pred[,1], probs = c(0.025,0.975))
quantile(y_pred[,2], probs = c(0.025,0.975))

```

#### Using Stan

Method 1: Using individual predictors

```stan
data {
  int<lower=0> n; // sample size
  vector[n] y;
  vector[n] x1;
  vector[n] x2;
  vector[n] x3;
  int<lower=0> n2; // sample size of the future observations
  vector[n2] x1_new;
  vector[n2] x2_new;
  vector[n2] x3_new;
}
parameters {
  real beta0;
  real beta1;
  real beta2;
  real beta3;
  real<lower=0> sigma2;
}
model {
  y ~ normal(beta0 + beta1 * x1 + beta2 * x2 + beta3 * x3, sqrt(sigma2)); // data distribution
  target += -log(sigma2); // prior
}
generated quantities {
  vector[n2] y_pred;
  for (i in 1:n2) {
    y_pred[i] = normal_rng(beta0 + beta1 * x1_new[i] + beta2 * x2_new[i] + beta3 * x3_new[i], sqrt(sigma2)); //the y values predicted by the model
  }
}
```

```{r}
#| cache: true

# specify stand code
stan_code1 <- "
data {
  int<lower=0> n;
  vector[n] y;
  vector[n] x1;
  vector[n] x2;
  vector[n] x3;
  int<lower=0> n2;
  vector[n2] x1_new;
  vector[n2] x2_new;
  vector[n2] x3_new;
}
parameters {
  real beta0;
  real beta1;
  real beta2;
  real beta3;
  real<lower=0> sigma2;
}
model {
  y ~ normal(beta0 + beta1 * x1 + beta2 * x2 + beta3 * x3, sqrt(sigma2));
  target += -log(sigma2);
}
generated quantities {
  vector[n2] y_pred;
  for (i in 1:n2) {
    y_pred[i] = normal_rng(beta0 + beta1 * x1_new[i] + beta2 * x2_new[i] + beta3 * x3_new[i], sqrt(sigma2));
  }
}
"

# save, provide needed info, and run model
stan_mod1 <- rstan::stan_model(model_code = stan_code1)
data_stan1 <- list(n = length(y), n2 = length(x1_new),
                   y = y, x1 = x1, x2 = x2, x3 = x3,
                   x1_new = x1_new, x2_new = x2_new, x3_new = x3_new)
stan_fit1 <- rstan::sampling(object = stan_mod1, data = data_stan1,
                             seed = 1, chains = 2, warmup = 2000, iter = 5000)

# view results
print(stan_fit1, digits = 2)

```

Method 2: Using the model matrix

```stan
data {
  int n; // the number of observations
  int n2; // the size of the new_X matrix
  int p; // the number of columns in the model matrix
  real y[n]; // the response
  matrix[n,p] X; // the model matrix
  matrix[n2,p] X_new; // the matrix for the predicted values
}
parameters {
  vector[p] beta; // the regression parameters
  real<lower=0> sigma2; // the variance
}
model {
  y ~ normal(X * beta, sqrt(sigma2));
  target += -log(sigma2);
}
generated quantities {
  vector[n2] y_pred;
  for (i in 1:n2) {
    y_pred[i] = normal_rng(X_new[i,] * beta, sqrt(sigma2));
  }
}
```

```{r}
#| cache: true

# specify stand code
stan_code2 <- "
data {
  int n;
  int n2;
  int p;
  real y[n];
  matrix[n,p] X;
  matrix[n2,p] X_new;
}
parameters {
  vector[p] beta;
  real<lower=0> sigma2;
}
model {
  y ~ normal(X * beta, sqrt(sigma2));
  target += -log(sigma2);
}
generated quantities {
  vector[n2] y_pred;
  for (i in 1:n2) {
    y_pred[i] = normal_rng(X_new[i,] * beta, sqrt(sigma2));
  }
}
"

# save, provide needed info, and run model
stan_mod2 <- rstan::stan_model(model_code = stan_code2)
data_stan2 <- list(n = length(y), n2 = length(x1_new), p = ncol(X),
                   y = y, X = X,
                   X_new = cbind(x0_new = 1, x1_new = x1_new, x2_new = x2_new, x3_new = x3_new))
stan_fit2 <- rstan::sampling(object = stan_mod2, data = data_stan2,
                             seed = 1, chains = 2, warmup = 2000, iter = 5000)

# view results
print(stan_fit2, digits = 2)

```

### Bayesian linear regression models: Analysis with conjugate priors

```stan
data {
  int n; // the number of observations
  int n2; // the size of the new_X matrix
  int p; // the number of columns in the model matrix
  real y[n]; // the response
  matrix[n,p] X; // the model matrix
  matrix[n2,p] X_new; // the matrix for the predicted values
}
parameters {
  vector[p] beta; // the regression parameters
  real<lower=0> sigma2; // variance
}
model {
  y ~ normal(X * beta, sqrt(sigma2));
  for (i in 1:p) {
    beta[i] ~ normal(0,100);
  }
  sigma2 ~ inv_gamma(0.01,0.01);
}
generated quantities {
  vector[n2] y_pred;
  for (i in 1:n2) {
    y_pred[i] = normal_rng(X_new[i,] * beta, sqrt(sigma2));
  }
}
```

```{r}
#| cache: true

# specify stand code
stan_code3 <- "
data {
  int n;
  int n2;
  int p;
  real y[n];
  matrix[n,p] X;
  matrix[n2,p] X_new;
}
parameters {
  vector[p] beta;
  real<lower=0> sigma2;
}
model {
  y ~ normal(X * beta, sqrt(sigma2));
  for (i in 1:p) {
    beta[i] ~ normal(0,100);
  }
  sigma2 ~ inv_gamma(0.01,0.01);
}
generated quantities {
  vector[n2] y_pred;
  for (i in 1:n2) {
    y_pred[i] = normal_rng(X_new[i,] * beta, sqrt(sigma2));
  }
}
"

# save, provide needed info, and run model
stan_mod3 <- rstan::stan_model(model_code = stan_code3)
data_stan3 <- list(n = length(y), n2 = length(x1_new), p = ncol(X),
                   y = y, X = X,
                   X_new = cbind(x0_new = 1, x1_new = x1_new, x2_new = x2_new, x3_new = x3_new))
stan_fit3 <- rstan::sampling(object = stan_mod3, data = data_stan3,
                             seed = 1, chains = 2, warmup = 2000, iter = 5000)

# view results
print(stan_fit3, digits = 2)

```

### Practice

Setup:

`data-haircut.csv` is a data set about the gender and cost of last haircut they paid for. Could you fit a model to decide if female student spend more on haircut on average? In this data set, some `NA`s are existing for the response (money spent on haircut). You need to remove these; additionally, the predictor variable `gender` needs to be recoded as 0 and 1 to fit the model.

Solution:

With gender as an indicator variable, the model statements become:


$$
\begin{align*}
y = \beta_0 + \beta_1 I_{Male} + \epsilon\\
\end{align*}
$$

$$
\begin{align*}
\Rightarrow y &= \beta_0 + \epsilon \quad \text{(Female)}\\
y &= \beta_0 + \beta_1 + \epsilon \quad \text{(Female)}\\
\end{align*}
$$

```stan
data {
  int n;
  int n2;
  int p;
  real y[n];
  matrix[n,p] X;
  matrix[n2,p] X_new;
}
parameters {
  vector[p] beta;
  real<lower=0> sigma2;
}
model {
   y ~ normal(X * beta, sqrt(sigma2));
   target += -log(sigma2);
   }
generated quantities {
  vector[n2] y_pred;
  for (i in 1:n2) {
    y_pred[i] = normal_rng(X_new[i,] * beta, sqrt(sigma2));
  }
}
```

```{r}
#| cache: true

# load and clean data
data_haircut <- read_csv(file = "data//data-haircut.csv") %>% 
  drop_na() %>%
  mutate(gender = ifelse(gender == "female", 0, 1))

# specify stand code
stan_code4 <- "
data {
  int n;
  int n2;
  int p;
  real y[n];
  matrix[n,p] X;
  matrix[n2,p] X_new;
}
parameters {
  vector[p] beta;
  real<lower=0> sigma2;
}
model {
   y ~ normal(X * beta, sqrt(sigma2));
   target += -log(sigma2);
   }
generated quantities {
  vector[n2] y_pred;
  for (i in 1:n2) {
    y_pred[i] = normal_rng(X_new[i,] * beta, sqrt(sigma2));
  }
}
"

# save, provide needed info, and run model
stan_mod4 <- rstan::stan_model(model_code = stan_code4)
x1_new <- c(0,1)
data_stan4 <- list(n = nrow(data_haircut), n2 = length(x1_new), p = (ncol(data_haircut) - 1) + 1,
                   y = data_haircut$haircut, X = cbind(x0 = 1, data_haircut$gender),
                   X_new = cbind(x0 = 1, x1_new = x1_new))
stan_fit4 <- rstan::sampling(object = stan_mod4, data = data_stan4,
                             seed = 1, chains = 2, warmup = 2000, iter = 5000)

# view results
print(stan_fit4, digits = 2)

# obtain posterior samples
list_draws <- rstan::extract(stan_fit4)
hist(list_draws$beta[,1])
hist(list_draws$beta[,2])
hist(list_draws$y_pred[,1])
hist(list_draws$y_pred[,2])

```

We see that the the estimated intercept averages `r round(mean(list_draws$beta[,1]), 2)` and the estimated $\beta_1$ coeﬃcient is around  `r round(mean(list_draws$beta[,2]), 2)`. This indicates that women tend to spend more on haircuts, on average, than men.
